---
title: "Homework 8"
subtitle:  "Significance Testing"
graphics: yes
output: pdf_document
header-includes:
    - \usepackage{amsmath,diagbox}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fastR2)
```

* * * 
### Instructions

Please answer the following questions in the order in which they are posed. Add a few empty lines below each and write your answers there. \textbf{Focus on answering in complete sentences and show work whether we ask for it or not}. You will also need scratch paper/pen to work out the answers before typing it.

For help with formatting documents in RMarkdown, please consult [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/). Another option is to search using Google. 

* * *

### Exercises 

<!---
I highly recommend that when you print the sampling distribution of a discrete test statistic, that you print both the value $x$ and the probability. For example, to print the Poisson probabilities, you might type something like the following. ---> 

```{r echo = F, eval=F}

pois_prob <- tibble(
     x = 0:10,
     f = dpois(x, lambda = 0.8)
)

pois_prob

```


1. (NHL) After a 2010 NHL playoff win in which Detroit Red-Wings wingman Henrik Zetterberg scored two goals in a 3-0 win over the Phoneix Coyotes. Detroit coach Mike Babcock said "He's been really good at playoff time each and every year. \emph{He seems to score at a higher rate in playoffs compared to the regular season."} 

   In 506 \textbf{regular season games}, Zetterberg scored 206 goals (goal scoring rate of $\frac{206}{506} = 0.407$). In 89 \textbf{playoff games}, he scored 44 goals (goal scoring rate of $\frac{44}{89} = 0.494$) Clearly, he has a higher goal scoring rate in the playoffs, but can it be explained by the vagaries of random chance? Or is the difference statistically significant?
   
   Let $X$ denote the number of goals he scores in the 89 playoff games.  Assume $$X \sim Pois(89 \times \lambda_0)$$ where $\lambda_0$ is his goal scoring rate per game during the playoffs.
   
   a. State the null and alternative hypothesis. 
   
   b. Calculate the P-value and summarize what you learn from it. Show your code for calculating the P-value. (You do not need to show the entire Poisson distribution)
   
   c. Suppose Coach Babcock had said "He's a really different player in the playoffs".  How would your answers to a and b change? Show your code for re-calculating the P-value.
   
   
2. (Sign test) Suppose $Y_1, Y_2, \dots, Y_{19}$ are independent random variables drawn from some distribution,  and we are interested in the
parameter $\pi_0 = P(Y < 0)$ More precisely, we want to test $H_0: \pi_0 = 0.4$ against $H_1: \pi_0 < 0.4$  at the 5% level.

   A reasonable test statistic would be $X$, the number of negative observations in the sample. That is, 
$$X = X_1 + X_2 +\dots + X_{19}$$
   where 
\begin{align*}
   X_i &= \left\{ \begin{array}{cc}
                    1 & Y_i < 0 \\
                    0 & otherwise \end{array} \right.
\end{align*}

   a. What is the sampling distribution of $X$ if the null hypothesis is true? Give a brief justification for your answer. 

   b. For what values of $X$ would you reject the null hypothesis? Support your answer showing code/output as necessary.

   c. What is the Type I error probability of your test? Support your answer showing code/output as necessary. 

   d. Calculate the Type II error probability when $\pi_0 = 0.2$. Support your answer showing code/output as necessary.


3. (Sign test again) Suppose in problem 2,you know that $Y$ is normally distributed with mean $\mu_0$ and (known) standard deviation $\sigma_0=1$.

   a. Re-state the null and alternative hypothesis from problem 2 in terms of claims about $\mu_0$. 
   
   b. We are still interested in performing the test at the 5% level, but we will now use the sample mean, $\bar{Y}$, as our statistic. What is the sampling distribution of $\bar{Y}$ if the null hypothesis is true? Give a brief justification for your answer.
   
   c. For what values of $\bar{Y}$ will you reject the null hypothesis? Support your answer showing code/output as necessary.
   
   d. What is the Type I error probability of your test? Briefly explain why.
   
   e. Calculate the Type II error probability when $\pi_0 = 0.2$. Support your answer showing code/output as necessary. (You will need to figure out what $\mu_0$ value to evaluate the Type II error probability at so it corresponds to a $\pi_0=0.2$.)

