---
title: "Homework 7"
subtitle:  "Interval Estimation"
graphics: yes
output: pdf_document
header-includes:
    - \usepackage{amsmath,diagbox}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fastR2)
```

* * * 
### Instructions

Please answer the following questions in the order in which they are posed. Add a few empty lines below each and write your answers there. \textbf{Focus on answering in complete sentences and show work whether we ask for it or not}. You will also need scratch paper/pen to work out the answers before typing it.

For help with formatting documents in RMarkdown, please consult [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/). Another option is to search using Google. 

* * *

### Exercises 



1. (Measurement error) Recall the pH-meter from Homework 6 which was known to give readings that were systematically higher or lower by a quantity $\delta_0$. In order to estimate $\delta_0$, six measurements $X_1, X_2, \dots, X_6$ were made from a solution with pH **known** to be 4.84. In your previous homework, you were asked to come up with an estimator for $\delta_0$. Let's call it $\hat{\delta}_0^{mom}$.

   Now, suppose four measurements  - $Y_1,Y_2,Y_3,Y_4$ - are made from a solution with an unknown pH-level $\mu_0$ resulting in 4.33, 4.22, 4.23, 4.37. As in the previous homework, the measurement error model is that $Y_1,Y_2,Y_3,Y_4$ is drawn independently from a distribution with mean $\mu_0 + \delta_0$ and variance $\sigma^2_0$. 

   Consider the estimator $$\hat{\mu}_0 = \bar{Y} - \hat{\delta}_0^{mom}$$ for $\mu_0$.

a. Show that $\hat{\mu}_0$ is an unbiased estimator of $\mu_0$. 

b. Give an expression for the standard error of $\hat{\mu}_0$. That is, find $\sqrt{ Var\left(\hat{\mu}_0 \right)}$. Show your work. (State any assumptions you need to make)

c. The variability in the pH measurements - $\sigma_0$ - is the same for both the $X$ measurements and also the $Y$ measurements. This makes sense since the variability in the readings is related to the meter, not the specific solution it is being used on. 

   A natural estimate for $\sigma_0$ is a pooled standard deviation $s_p$ calculated from both samples. The formula for $s_p$ is below: 
$$s^2_p = \frac{ \sum\limits_{i=1}^{6} (x_i - \bar{x})^2 + \sum\limits_{j=1}^{4} (y_i - \bar{y})^2}{6+4-2}$$

   Calculate $s_p$, the pooled estimate of $\sigma_0$. 

```{r est-sigma}
#six measurements for solution with pH = 4.84 from homework 6
x<- c(4.71, 4.63, 4.69, 4.76, 4.58, 4.83)   
#four measurements for solution with unknown pH from this homework
y<- c(4.33, 4.22, 4.23, 4.37)
```


d. Calculate the estimated standard error of $\hat{\mu}_0$. Show your steps.


2. (Force) A type of metal bar breaks when a force of size $X$ is applied, where $X$ has PDF
$$f(x) = 2 \alpha_0 x\: e^{-\alpha_0\:x^2} \ \ \ \ x > 0$$
where $\alpha_0 > 0$ is an unknown parameter. 
We observe a breaking force of 40. Find a 95% confidence interval for $\alpha_0$.

   Hint: We are looking for a random interval $[L, U]$ which contains $\alpha_0$ with probability 95%. Construct the interval by "inverting" the probability statement
   $$P\left( q_{0.025} \leq X \leq q_{0.975} \right) = 0.95$$
   where $q_{0.025}$ and $q_{0.975}$ are the 2.5th and 97.5th percentiles of the distribution of $X$.
   
3. (CLT) A sample of 83 observations for an integer-valued random variable $Y$ is shown below:
\begin{table}[h]
\centering
\begin{tabular}{ccccccc} \\ \hline
value & 0 & 1 & 2 & 3 & 4 & 5 \\ \hline
frequency & 13 & 18 & 23 & 15 & 6 & 8 \\ \hline
\end{tabular}
\end{table}

   Use the Central Limit Theorem to find a 90% confidence interval for $\pi_0 = P(Y \geq 2)$. Show your work, develop your answer. We are grading on style.

   Hint: You actually have 83 independent Bernoulli random variables - $X_1,X_2,\dots,X_{83}$ - where each $X_i$ is one if $Y \geq 2$ and zero otherwise. Therefore you can think of $X_1, X_2, \dots, X_{83} \overset{i.i.d.}{\sim} Binom(1, \pi_0)$ and you wish to construct a confidence interval for the mean of the distribution - $\pi_0$ -  using the CLT.


4. (Airbnb) Read sections 18.3 and 19.2 in the Notes where I constructed a confidence interval for the mean (daily) price of 2 bedroom apartment rentals in Seattle. In this section you will repeat this calculation for a different subset of rentals: houses with 3 or more bedrooms where the entire home is for rent. The variables you will be filtering on and their values are shown below: 
    - property_type: Houses
    - room_type: Entire home/apt
    - bedrooms: 3 or more

a. In this part, you will construct a large sample 95% confidence interval for the mean price of all such house rentals in Seattle. 
   Be sure to 
   
   - display the first five rows of the filtered data frame (showing just price)
   
   - make a histogram of `price` and 
   
   - calculate and report a large sample 95% confidence interval for the mean daily price. (See section 18.3 from pages 206-208 for example code.)
   
b. In this part, you will construct a (non-parametric) bootstrap confidence interval for the mean price of houses with 3 or more bedrooms where the entire home is for rent. Be sure to

   - display the bootstrap sampling distribution of the sample mean

   - compare the bootstrap sampling distribution with the normal distribution 

   - calculate and report the standard bootstrap confidence limits
   (See section 19.2 on pages 219 - 222 for example code)
   
```{r boot}
set.seed(14141)
B = 1000

```
  
  