---
title: "Problem Section 1 Key"
subtitle:  "Joint, Marginal and Conditional"
graphics: yes
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* * * 
### Instructions

Please answer the following questions in the order in which they are posed. Add a few empty lines below each and write your answers there. Focus on answering in complete sentences. You will also need scratch paper/pen to work out the answers before typing it.

For help with formatting documents in RMarkdown, please consult [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/). Another option is to search using Google. 

* * *

### Exercises 

1. A fair coin is tossed four times. $X$ is the number of heads that come up on the first three tosses and $Y$ is the number of heads that come up on tosses 2, 3, 4. 

   a. Make a table showing the joint PMF of $\langle X, Y \rangle$. 
   
   \begin{tabular}{|c|c|c|c|c|}
   \hline
    & X=0 & X=1 & X=2 & X=3 \\
   \hline
   Y=0 & `r .5^4` & `r .5^3*.5` & 0 & 0 \\
   Y=1 & `r .5^3*.5` & `r .5^4 + .5^2*.5^2*2` & `r .5*.5^2*2*.5` & 0 \\
   Y=2 & 0 & `r .5*.5^2*2*.5` & `r .5^4 + .5^2*.5^2*2` & `r .5^4` \\
   Y=3 & 0 & 0 & `r .5^3*.5` & `r .5^4` \\
   \hline
   \end{tabular}
   
   By the multiplication rule there are $$2 \times 2 \times 2 \times 2 = 2^{4} = 16$$ possible outcomes in this sample space. Each outcome is equally likely and therefore the probabilities of the various $(x,y)$ combinations can be deduced by simply counting the number of outcomes that would result in this value.
   
   For example, the outcomes that lead to $x=1, y=2$ are HTTH, THTH, TTHH, which gives a probability of 3/16.

   
   b. Find the marginal distribution of $X$ and also of $Y$.
   
   We can find the marginal distributions by simply adding the probabilities along the rows or the columns. However, we can also recognize that since $X$ is the number of heads in 3 tosses, it is a binomial random variable. Hence it's probabilities are given by the $Binom(3, \frac{1}{2})$ distribution.
   
   \begin{tabular}{|c|c|c|c|c|}
   \hline
    & X=0 & X=1 & X=2 & X=3 \\
    \hline
   $f_x(x)$ & `r dbinom(0,3,.5)` & `r dbinom(1,3,.5)` & `r dbinom(2,3,.5)` & `r dbinom(3,3,.5)` \\
   \hline
   \end{tabular}
   
   \begin{tabular}{|c|c|c|c|c|}
   \hline
    & Y=0 & Y=1 & Y=2 & Y=3 \\
    \hline
   $f_y(y)$ & `r dbinom(0,3,.5)` & `r dbinom(1,3,.5)` & `r dbinom(2,3,.5)` & `r dbinom(3,3,.5)` \\
   \hline
   \end{tabular}
   
   
   c. Determine the conditional distribution $f(x|y)$ for $y = 2$. 
   
$$
f(x|y=2) = \frac{f(x,y=2)}{f_y(2)} = \frac{f(x,y=2)}{0.375}
$$
Thus using the values from our table in part a we have:

\begin{tabular}{|c|c|c|c|c|}
   \hline
    & X=0 & X=1 & X=2 & X=3 \\
    \hline
   $f(x|y=2)$ & 0 & `r (.5*.5^2*2*.5)/.375` & `r (.5^4 + .5^2*.5^2*2)/.375` & `r .5^4/.375` \\
   \hline
   \end{tabular}


2. A physics professor has just given an exam to fifty students enrolled in a thermodynamics class. From past experience, she has reason to believe that the scores will be
normally distributed with $\mu = 80.0$ and $\sigma = 5.0$. Students scoring ninety or above will
receive A’s, between eighty and eighty-nine, B’s, and so on. What are the expected
values and variances for the numbers of students receiving each of the five letter
grades? Support your answer.

First we will denote the probability for falling into each grade bucket using the probabilities calculated from a normal distribution. We will assume that the scores are unrestricted and can extend to $\pm \infty$ when doing the calculations. 

We have:

$\pi_a$ = `r pnorm(90,80,5,lower.tail=F)`
$\pi_b$ = `r -diff(pnorm(c(90,80),80,5))`
$\pi_c$ = `r -diff(pnorm(c(80,70),80,5))`
$\pi_d$ = `r -diff(pnorm(c(70,60),80,5))`
$\pi_f$ = `r pnorm(60,80,5,lower.tail=T)`

```{r}
pi_a = pnorm(90,80,5,lower.tail=F)
pi_b = -diff(pnorm(c(90,80),80,5))
pi_c = -diff(pnorm(c(80,70),80,5))
pi_d = -diff(pnorm(c(70,60),80,5))
pi_f = pnorm(60,80,5,lower.tail=T)
```


Denote the random variables $X_A, X_B, X_C, X_D, X_F$ which correspond to the number of students who received each of those letter grades in the class of 50. We thus have that:

$$
(X_A, X_B, X_C, X_D, X_F) \sim Multinom(n=50,k=5,\pi = (\pi_a,\pi_b,\pi_c,\pi_d,\pi_f))
$$

Now using the properties of the multinomial distribution we have:
$$X_i \sim Binom(n, \pi_i).$$ (Ref: Thm 14.1)
Therefore

$E(X_i) = n\pi_i$
$Var(X_i) = n\pi_i(1-\pi_i)$

Using our values for $\pi$ above, we can easily calculate these values. 

```{r}
#Calculate the real values here
E_X_a <- 50*pi_a
var_X_a <- 50*pi_a*(1-pi_a)
```

We should thus expect `r E_X_a` number of students who receives an A, give or take `r sqrt(var_X_a)`


3. Suppose $X$ and $Y$ are discrete random variables with joint PMF given by the formula
$$f(x,y) = \frac{4!}{x! \:y! \:(4-x-y)!} \left(\frac{1}{2} \right)^{x} \left(\frac{1}{3} \right)^{y} \left(\frac{1}{6} \right)^{4-x-y}$$
where $x$, $y$ take values $0,1,2,3,4$ but are constrained by the requirement that $x+y \leq 4$.

a. Show that  
$$\frac{4!}{x! \:y! \:(4-x-y)!} = \binom{4}{x} \cdot \binom{4-x}{y}.$$
(This equality should help us understand why $\frac{4!}{x! \:y! \:(4-x-y)!}$ counts the number of outcomes which have $x$ A's and $y$ B's in a trinomial experiment with 4 independent trials.)

We may show the right hand side (RHS) is the same as the LHS by using the definition of the binomial function: 

$$
\binom{4}{x} \cdot \binom{4-x}{y} = \frac{4!}{x!(4-x)!}\frac{(4-x)!}{y!(4-x-y)!} = \frac{4!}{x!y!(4-x-y)!}
$$

b. We showed in class that $X \sim Binom(4, \frac{1}{2})$. Show that the conditional distribution of $Y$ given $X=x$ is also in the binomial family. Specifically show that 
$$f(y | x) = \binom{4-x}{y} \left(\frac{2}{3} \right)^{y} \left(\frac{1}{3} \right)^{4 - x - y}, \ \ y = 0, 1, 2, \dots, ,4-x.$$

We know that:

$$
f(y|x) = \frac{f(x,y)}{f(x)}
$$

Thus we may plug in the distributions and proceed with simplifying.

$$
f(y|x) = \frac{\frac{4!}{x! \:y! \:(4-x-y)!} \left(\frac{1}{2} \right)^{x} \left(\frac{1}{3} \right)^{y} \left(\frac{1}{6} \right)^{4-x-y}}{\binom{4}{x}(\frac{1}{2})^x(\frac{1}{2})^{4-x}}
$$

Using part b we know we can rewrite this as:

$$
f(y|x) = \frac{\binom{4}{x} \cdot \binom{4-x}{y} \left(\frac{1}{2} \right)^{x} \left(\frac{1}{3} \right)^{y} \left(\frac{1}{6} \right)^{4-x-y}}{\binom{4}{x}(\frac{1}{2})^x(\frac{1}{2})^{4-x}}
$$

After some cancelling out terms on the top and bottom we have:

\begin{align*}
f(y|x) &= \frac{\binom{4-x}{y}  \left(\frac{1}{3} \right)^{y} \left(\frac{1}{6} \right)^{4-x-y}}{(\frac{1}{2})^{4-x}} \\
&= \frac{\binom{4-x}{y}  \left(\frac{1}{3} \right)^{y} \left(\frac{1}{2} \right)^{4-x-y}\left(\frac{1}{3} \right)^{4-x-y}}{(\frac{1}{2})^{4-x}} \\
&= \binom{4-x}{y}  \left(\frac{1}{3} \right)^{y} \left(\frac{1}{2} \right)^{-y}\left(\frac{1}{3} \right)^{4-x-y}\\
&= \binom{4-x}{y}  \left(\frac{2}{3} \right)^{y}\left(\frac{1}{3} \right)^{4-x-y}
\end{align*}

Where this mass function is valid for values $y=0,1,2,\dots, 4-x$

