---
title: "Problem Section 6 KEY"
subtitle:  "Bootstrap Confidence Intervals"
graphics: yes
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

* * * 
### Instructions

Please answer the following questions in the order in which they are posed. Add a few empty lines below each and write your answers there. Focus on answering in complete sentences. You will also need scratch paper/pen to work out the answers before typing it.

For help with formatting documents in RMarkdown, please consult [R Markdown: The Definitive Guide](https://bookdown.org/yihui/rmarkdown/). Another option is to search using Google. 

* * *

### Exercises 


1. (Non-parametric bootstrap) Consider the set of 20 numbers which were independently drawn from some distribution with mean $\mu_0$ and standard deviation $\sigma_0$. 

```{r}
options(pillar.sigfig = 8)
sample_data <- tibble(
  x = c(3.56, 0.69, 0.10, 1.84, 3.93, 1.25, 0.18, 1.13, 0.27, 1.21, 
        0.50, 0.67, 0.01, 0.61, 0.82, 1.70, 0.39, 0.11, 1.20, 0.72)
)
n <- nrow(sample_data)
obs_sigma2_hat <-  var(sample_data$x)*(n-1)/n
obs_sigma2_hat
```

The observed value for the estimator of the population variance 
$$\hat{\sigma}^2 = \frac{1}{20} \sum\limits_{i=1}^{20} (x_i - \bar{x})^2$$
is 1.0668. 

We are interested in constructing the sampling distribution of this statistic  as an estimator for the population variance $\sigma^2_0$. In this problem you will create this using non-parametric bootstrap to resample from the original data. You will then calculate a confidence interval for $\sigma^2_0$.


```{r boot-s2}
set.seed(141)
B = 1000
n <- nrow(sample_data)

#1. Take B random samples of size 20 each from your data with replacement and calculate the value...
#...of s2 for the resample.
#2. Make a histogram of the bootstrapped values of s2. 
#3. Examine the shape, calculate the mean and standard deviation of the bootstrapped values.
#4. Calculate a 95% bootstrap confidence interval for sigma2 using the standard bootstrap... 
#...confidence interval and also the simple percentile method

#Step 1:
boot_df <- tibble(
  #calculate variance of each resample created using replicate
  var_star = replicate(n = B,
                       expr = var(sample(x = sample_data$x,
                                         size = nrow(sample_data),
                                         replace=TRUE) )*(n-1)/n) )
#Our DF will have a total of B variances
boot_df %>% slice_head(n=5)
```
```{r, message=F, warning=F}
#Step 2: Create a histogram

ggplot(data=boot_df, 
       mapping = aes(x=var_star,y=..density..))+
  geom_histogram(binwidth = .2)+
  labs(title="Histogram of 1000 Bootstrapped Sample Variance Values")
```

Step 3: Analyze the bootstrap samples

Here we see that this is a right skewed distribution.

For some summary statistics we see:

```{r}
boot_df_summary <- boot_df %>% summarize(bootstrap_mean_var = mean(var_star), 
                                         bootstrap_se = sd(var_star))
boot_df_summary
```


We see that the mean of the bootstrap variances, 1.01, is slightly different than the sample value we saw before from the raw sample. To calculate a confidence interval we can consider two different methods. 

In the first method we will use the standard bootstrap method.

First we can examine the normality of our samples.

```{r}
ggplot(data=boot_df,
       mapping = aes(sample = var_star))+
  stat_qq(distribution = qnorm)+
  stat_qq_line(distribution = qnorm)+
  labs(title = "Normal Prob. Plot of Bootstrap Variances")
```

We see that this is not very normal, especially at the left tail. We can proceed anyways with this standard method and examine the CI of this form. 

We must also do a bias correction here. We see that our bootstrap returned a mean variance of 1.011, while the original variance estimator yielded 1.0667. Thus we have a bias of `r (1.0109 - 1.0667)`. This will give us a bias corrected estimate of the population variance as:

```{r}
#sample - bias
1.0667 - (1.0109 - 1.0667)
```

That is we will have that our 95% CI will be:

$$
[(\hat{\sigma}^2 - bias) - 1.96*se_{bootstrap}, (\hat{\sigma}^2 - bias) + 1.96*se_{bootstrap}] = [1.12-1.96*.47,1.12+1.96*.47 ]
$$
Calculating this with the non-rounded numbers we see we have the interval as:

```{r}
c(obs_sigma2_hat - (boot_df_summary$bootstrap_mean_var - obs_sigma2_hat) + 
    c(-1,1)*qnorm(.975)*boot_df_summary$bootstrap_se)
```

We can also consider the simple percentile method. In this case we will simply take the 2.5 and 97.5 percentiles of our bootstrap samples. This yields:

```{r}
boot_df %>% summarize(lower = quantile(var_star,.025),
                      upper = quantile(var_star, .975))
```

In this case we would prefer the simple percentile method, as from looking at our qqplot, the data is not very normal so using normal quantiles does not seem very reasonable. 
